SMT-Based Context-Bounded Model Checking for Embedded Systems: Challenges and Future Trends

### 2016-03-25 paper.tex
remove et al.

Bounded Model Checking (BMC) based on Boolean Satisfiability (SAT) was originally proposed by Biere {\it et al.} to verify hardware designs~\cite{Biere99,handbook09}. The authors were able to successfully verify large digital circuits with approximately $9510$ latches and $9499$ inputs, leading to BMC formulae with $4 \times 10^6$ variables and $1.2 \times 10^7$ clauses to be checked by a standard SAT solver. BMC based on Satisfiability Modulo Theories (SMT)~\cite{BarrettSST09} was originally proposed by Armando {\it et al.}~\cite{Armando06} to deal with increasing software verification complexity.

@@

Gadelha {\it et al.}, Beyer {\it et al.}, and Brain {\it et al.} presented new verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant)~\cite{Gadelha15,Beyer15,Brain15,Rocha15}. Additionally, \textit{k}-induction based verification was also applied to ensure that (restricted) C programs (1) do not contain violations related to data races~\cite{Donaldson10,Kinductor} considering the  Cell BE processor and (2) do respect time constraints, which are specified during the system design phase~\cite{EenS03}. The \textit{k}-induction algorithm is also a well-established technique in hardware verification, where it is easily applied, due to the monolithic transition relation present in such designs~\cite{EenS03,Sheera00,GrosseLD09}.

@@

Indeed, research in software verification is now taking such considerations into account and some schemes already use knowledge about the system to be verified and the underlying hardware. Recently, Ismail {\it et al.} \cite{dsv_spin2015}  proposed a verification tool for digital systems, called digital system verifier (DSVerifier), which is able to aid engineers to check overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length effects. Indeed, DSVerifier is a useful test tool, which takes into account realization forms ({\it e.g.}, direct forms, transposed forms, etc.) and other implementation restrictions to explore the design space. Ultimately, if the system requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup.

@@

\textbf{(RP1)} Cordeiro {\it et al.} proposed the first SMT-based BMC for full C programs, called Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{Cordeiro12}, which was later extended to support C++98 programs~\cite{ECBS13}, CUDA programs~\cite{Pereira15}, and Qt-based consumer electronics applications~\cite{Sousa15}. This approach was also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were later confirmed by the benchmarks' creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC)~\cite{CordeiroF11,Cordeiro12}. Other SMT-based BMC approaches have also been proposed and implemented~\cite{MerzFS12}, but the coverage and verification time of all existing ones are still limited to specific classes of programs, especially for those that contain intensive floating-point arithmetic and dynamic memory allocation~\cite{Beyer14,BeyerSVCOMP15}. One possible research direction is to bridge the gap between BMC tools and SMT solvers to develop more efficient decision procedures to handle specific class of programs.

\textbf{(RP2)} The SMT-based BMC approach proposed by Cordeiro {\it et al}. was further developed to verify correct lock acquisition ordering and the absence of deadlocks, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering monotonic partial-order reduction~\cite{KahlonWG09} and state-hashing techniques, in order to prune the state-space exploration~\cite{morse15}. Recent advances for verifying multi-threaded C programs have been proposed to speed up the verification time, which significantly prune the state-space exploration~\cite{Inverso14,civl15}. However, the class of concurrent programs ({\it e.g.}, CUDA, OpenCL, and MPI) that can be verified is still very limited. One possible research direction is to further extend BMC of multi-threaded C programs via Lazy Sequentialization~\cite{Inverso14}, in order to analyze unsatisfiability cores with the goal to remove redundant behaviour.

>>>>

Bounded Model Checking (BMC) based on Boolean Satisfiability (SAT) was originally proposed to verify hardware designs~\cite{Biere99,handbook09}. The authors were able to successfully verify large digital circuits with approximately $9510$ latches and $9499$ inputs, leading to BMC formulae with $4 \times 10^6$ variables and $1.2 \times 10^7$ clauses to be checked by a standard SAT solver. BMC based on Satisfiability Modulo Theories (SMT)~\cite{BarrettSST09} was originally proposed to deal with increasing software verification complexity~\cite{Armando06}.

@@

Novel verification algorithms for proving correctness of (a large set of) C programs, by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant) were proposed~\cite{Gadelha15,Beyer15,Brain15,Rocha15}. Additionally, \textit{k}-induction based verification was also applied to ensure that (restricted) C programs (1) do not contain violations related to data races~\cite{Donaldson10,Kinductor} considering the  Cell BE processor and (2) do respect time constraints, which are specified during the system design phase~\cite{EenS03}. The \textit{k}-induction algorithm is also a well-established technique in hardware verification, where it is easily applied, due to the monolithic transition relation present in such designs~\cite{EenS03,Sheera00,GrosseLD09}.

@@

Indeed, research in software verification is now taking such considerations into account and some schemes already use knowledge about the system to be verified and the underlying hardware. Recently, a verification tool proposed for digital systems, called digital system verifier (DSVerifier)~\cite{dsv_spin2015}, which is able to aid engineers to check overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length effects. Indeed, DSVerifier is a useful test tool, which takes into account realization forms ({\it e.g.}, direct forms, transposed forms, etc.) and other implementation restrictions to explore the design space. Ultimately, if the system requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup.

@@

\textbf{(RP1)} Cordeiro, Fischer, and Marques-Silva proposed the first SMT-based BMC for full C programs, called Efficient SMT-Based Context-Bounded Model Checker (ESBMC)~\cite{Cordeiro12}, which was later extended to support C++98 programs~\cite{ECBS13}, CUDA programs~\cite{Pereira15}, and Qt-based consumer electronics applications~\cite{Sousa15}. This approach was also able to find undiscovered bugs related to arithmetic overflow, buffer overflow, and invalid pointer, in standard benchmarks, which were later confirmed by the benchmarks' creators ({\it e.g.}, NOKIA, NEC, NXP, and VERISEC)~\cite{CordeiroF11,Cordeiro12}. Other SMT-based BMC approaches have also been proposed and implemented~\cite{MerzFS12}, but the coverage and verification time of all existing ones are still limited to specific classes of programs, especially for those that contain intensive floating-point arithmetic and dynamic memory allocation~\cite{Beyer14,BeyerSVCOMP15}. One possible research direction is to bridge the gap between BMC tools and SMT solvers to develop more efficient decision procedures to handle specific class of programs.

\textbf{(RP2)} The SMT-based BMC approach proposed by Cordeiro and Fischer was further developed to verify correct lock acquisition ordering and the absence of deadlocks, data races, and atomicity violations in multi-threaded software based on POSIX and CUDA libraries~\cite{CordeiroF11,Pereira15}, considering monotonic partial-order reduction~\cite{KahlonWG09} and state-hashing techniques, in order to prune the state-space exploration~\cite{morse15}. Recent advances for verifying multi-threaded C programs have been proposed to speed up the verification time, which significantly prune the state-space exploration~\cite{Inverso14,civl15}. However, the class of concurrent programs ({\it e.g.}, CUDA, OpenCL, and MPI) that can be verified is still very limited. One possible research direction is to further extend BMC of multi-threaded C programs via Lazy Sequentialization~\cite{Inverso14}, in order to analyze unsatisfiability cores with the goal to remove redundant behaviour.
